---
title: "Inferring Ornstein-Uhlenbeck hyperparameters"
author: "J.P. Meagher"
date: "12 October 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F)
```

Now that the weights associated with each hyperparameter have been identified the hyperparameters for each of the independent phylogenetic Ornstein-Uhlenbeck processes need to be inferred.

# Packages

```{r packages}
library(batwork)
library(sdsBAT)
library(ape)
library(tidyverse)
library(magrittr)
library(RColorBrewer)
```

# Data

```{r data}
observed_frequencies <- get_spectrogram_details(
  mexican_bat_calls %>% use_series('call') %>% extract2(1), 
  detail = 'freq')

band <- observed_frequencies > 9000 & observed_frequencies < 212000

df <- mexican_bat_calls %>% 
  mutate(time = map(calls, get_spectrogram_details, detail = 'time'), 
    psd = map(calls, get_psd)) %>% 
  mutate(restricted = map(psd, extract, , band)) %>% 
  mutate(restricted = map(restricted, log10)) %>% 
  mutate(restricted = map(restricted, multiply_by, 10)) %>% 
  mutate(smoothed = map(restricted, waveslim::denoise.dwt.2d, J = 2)) %>% 
  mutate(regularised = map(smoothed, grid_interpolation)) %>%
  select(bat, species, sex, regularised)

f <- seq(observed_frequencies %>% extract(band) %>% min,
  observed_frequencies %>% extract(band) %>% max,
  length.out = df %>% use_series('regularised') %>% extract2(1) %>% ncol)

t <- seq(0, 1,
  length.out = df %>% use_series('regularised') %>% extract2(1) %>% nrow)

df <- df %>% 
  mutate(vectorised = map(regularised, c))

mean_spectrogram <- df %>%  
  select(vectorised) %>%
  unlist %>% 
  matrix(nrow = 50*100, ncol = df %>% nrow) %>% 
  rowMeans

df <- df %>% 
  mutate(centred = map(vectorised, subtract, mean_spectrogram))

components <- df %>% 
  select(regularised) %>%
  unnest %>% 
  use_series('regularised') %>% 
  array(dim = c(50*100, df %>% nrow)) %>% 
  t %>% 
  prcomp %>% 
  use_series(rotation)
component_weights <- function(centred_spectrogram, evolutionary_components){
  weights <- lm(centred_spectrogram ~ evolutionary_components) %>% use_series(coefficients)
  return(weights)
}

df <- df %>% 
  mutate(weights = map(centred, component_weights, evolutionary_components = components[, 1:6]))

wt <- df %>% 
  select(bat, sex, species, weights) %>% 
  mutate(weights = map(weights, matrix, nrow = 1)) %>% 
  mutate(weights = map(weights, as.tibble)) %>% 
  unnest
```

Having obtained the weight of each component associated with each spectrogram a resampling scheme can be used to infer the most likely hyperparameter values.

For this I am going to try to use all the available samples. This means constructing a new phylogenetic tree with the appropriate number of tips.

```{r}
n_samples <- df %>% group_by(species) %>% summarise(n_recordings = species %>% length)

tree <- phylogeny
expanded_tree <- phylogeny
i <- tree %>% use_series(tip.label) %>% extract(1)

for(i in tree %>% use_series(tip.label)){
  ref <- n_samples %>% filter(species == i)
  
  individuals <- stree(ref %>% use_series(n_recordings))
  individuals$edge.length <- rep(0, ref %>% use_series(n_recordings))
  individuals$tip.label <- rep(i, ref %>% use_series(n_recordings))
  
  expanded_tree <- bind.tree(expanded_tree, individuals, which(expanded_tree$tip.label == i))
}
```

Given the full pylogenetic tree for the sample the observations of weight can be passed to a type 2 maximum likelihood estimator for the kernel hyperparameters.

```{r}
expanded_tree %>% use_series(tip.label) %>% length

hyper <- matrix(nrow = 4, ncol = 6)
j <- 1
for(i in 5:10){
  comp <- wt %>% extract(,i)
  names(comp) <- wt %>% use_series(species)
  comp <- comp[expanded_tree %>% use_series(tip.label)]
  
  h <- sdsBAT:::pou_type2mle(comp, expanded_tree, pou_logl_fast, upper_initialisation = c(10,100,10))
  hyper[,j] <- h
  j <-  j+1
}

rownames(hyper) <- c('s_p', 'l', 's_n', 'lik')
hyper
```

It seems that the phylogenetic variance is very high, and the length scales seem to be a little shorter than ideal.

But we are getting some reasonably interesting results.