---
title: "Independent Components Analysis"
author: "J.P. Meagher"
date: "13 October 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F)
```

Evolutionary Components have been approximated by Principal components. In this analysis I will instead find independent components wich may provide more satisfying components.

# Packages

```{r packages}
library(batwork)
library(signal)
library(sdsBAT)
library(ape)
library(tidyverse)
library(magrittr)
library(RColorBrewer)
```

# Data

```{r data}
observed_frequencies <- get_spectrogram_details(
  mexican_bat_calls %>% use_series('call') %>% extract2(1), 
  detail = 'freq')

band <- observed_frequencies > 9000 & observed_frequencies < 212000

df <- mexican_bat_calls %>% 
  mutate(time = map(calls, get_spectrogram_details, detail = 'time'), 
    psd = map(calls, get_psd)) %>% 
  mutate(restricted = map(psd, extract, , band)) %>% 
  mutate(restricted = map(restricted, log10)) %>% 
  mutate(restricted = map(restricted, multiply_by, 10)) %>% 
  mutate(smoothed = map(restricted, waveslim::denoise.dwt.2d, J = 2)) %>% 
  mutate(regularised = map(smoothed, grid_interpolation)) %>%
  select(bat, species, sex, regularised)

f <- seq(observed_frequencies %>% extract(band) %>% min,
  observed_frequencies %>% extract(band) %>% max,
  length.out = df %>% use_series('regularised') %>% extract2(1) %>% ncol)

t <- seq(0, 1,
  length.out = df %>% use_series('regularised') %>% extract2(1) %>% nrow)

components <- df %>% 
  select(regularised) %>%
  unnest %>% 
  use_series('regularised') %>% 
  array(dim = c(50*100, df %>% nrow)) %>% 
  t %>% 
  prcomp %>% 
  use_series(rotation)
```

# Orthogonal Components

First check that the principal components are indeed orthoghonal.

```{r}
orth <- (t(components) %*% components)

orth %>% image
```

Principal components ar indeed orthogonal.

These orthogonal. This will be done by cubica algorithm.

```{r}
ica <- cubica34(components %>% extract(, 1:6) %>% t)

(ica %>% use_series(y)) %*% (t(ica %>% use_series(y)))
```

# Visual Inspection

Components in this case will need to be sense checked visually.

```{r}
n <- 1
while(n <= 6){
  components %>% 
  extract(,n) %>% 
  matrix(nrow = 50, ncol = 100) %>% 
  image(t, f/1000, .,  main = paste('Principal Component', n), 
  xlab = 'Time', ylab = 'Frequency (kHz)',
  col = brewer.pal(9, 'Spectral'))

ica %>% 
  use_series(y)%>% 
  extract(n,) %>% 
  matrix(nrow = 50, ncol = 100) %>% 
  image(t, f/1000, .,  main = paste('Independent Component', n), 
  xlab = 'Time', ylab = 'Frequency (kHz)',
  col = brewer.pal(9, 'Spectral'))
n <- n + 1
}
```

