---
title: "Untitled"
author: "J.P. Meagher"
date: "29 November 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F)
```

This study considers Principal Components as Evolutionary Features and investigates the quality of reconstructions taking the mean absolute error between actual and reconstructed echolocation call spectrograms. This will then be investigated in the context of a LOSO-CV of phylogenetic reconstructions of call spectrograms.

```{r packages}
library(signal) # Produce Spectrogram
library(tidyverse) # see http://r4ds.had.co.nz/
library(magrittr) # pipe operator and alaises
library(batwork) # my own package, long form call data
library(sdsBAT) # my own package, tree data and functions for ancestral reconstruction
library(ape) # Analyses of Phylogenetics and Evolution
library(RColorBrewer) # for pretty pictures
library(ggridges) # Joy plots
library(ggtree) ## ggplot for phylogenetic trees
```

# Spectrogram Inversion Stuff

```{r complex stft, echo = T}
complex_stft <- function(x, window_size = min(256, length(x)), 
  window_function = hamming(window_size), step_size = ceiling(window_size/2)){
  if (!is.numeric(x)) stop("'x' has to be a numeric.")
  
  if (length(window_size) > 1) stop("Cannot comput for multiple window sizes")
  
  if (length(x) > window_size) {
    window_offset <- seq(1, length(x) - window_size, by = step_size)
  }else{
    window_offset <- 1
    }
  
  sections <- matrix(nrow = window_size, ncol = length(window_offset))
  for(i in seq_along(window_offset)){
    sections[, i] <- 
      x[window_offset[i]:(window_offset[i] + window_size - 1)]*window_function
  }

  stft <- mvfft(sections)
  
  return(stft)
}
```

## Observation Positions: Frequency Domain

Given observations over a spectrum such as that produced by the `complex_stft` function, we want to be able to index the frequencies at which the intensity has been calculated. The order in which frequencies are calculated by the discrete Fourier transform is $\{0, 1,\dots, \frac{n}{2} - 1, \pm \frac{n}{2}, -(\frac{n}{2} - 1), \dots, -2,-1\}$ for series with an even number of observations and $\{0, 1,\dots, \frac{n - 1}{2}, -\frac{n - 1}{2}, -(\frac{n}{2} - 1), \dots, -2,-1\}$ for those with an odd number of observations. The `full_spectrum_frequencies` combines this information with the sampling rate to find the frequency in Hz to wich our spectrogram frequencies correspond.

```{r full spectrum frequencies, echo = T}
full_spectrum_frequencies <- function(full_spectrum, fs = 500){
  if(is.vector(full_spectrum)){
    
    n_freq <- length(full_spectrum)
    
  }else if(is.matrix(full_spectrum)){
      
    n_freq <- nrow(full_spectrum)
    
  }else{
    
      stop('full_spectrum is in incorrect format')
  }
  
  
  if(n_freq %% 2 == 1){
    frequencies <- seq(from = 0, to = floor(n_freq/2), length.out = (n_freq + 1)/2)
    frequencies <- c( frequencies, -rev( frequencies[ 2:((n_freq + 1)/2) ] ) )
  }else{
    frequencies <- seq(from = 0, to = n_freq/2, length.out = (n_freq/2) + 1)
    frequencies <- c( frequencies, -rev( frequencies[ 2:(n_freq/2) ] ) )
  }
  
  frequencies <- frequencies * (fs/ n_freq)
  
  return(frequencies)
}
```

## Observation Positions: Time Domain

In the STFT, windows are centred at a particular point in the time of the original signal. These centre points of each window are calculated by the `stft_time` function.

```{r stft time, echo = T}
stft_time <- function(stft, fs, step_size){
  
  if(!is.matrix(stft)){
    stop('spectrum must be a matrix')
  }
  
  window_offset <- seq(1, by = step_size, length.out = ncol(stft))
  times <- window_offset / fs
  
  return(times)
}
```

## Magnitude STFT

Complications in inverting a Magnitude STFT arise as phase information about the signal has been discarded. This is where RTISI-LA algorithm will become useful. Firstly however, simply consider the structure of phase information.

Given the phase associated with a magnitude stft, the inversion is straightforward.

The process becomes more complicated when there is no available phase information. In this case we generate a random phase for the first window of the signal and then try to use this to infer the phase of the next frames.

```{r rtisi-la, echo = F}
random_phase <- function(window_size){
  if(window_size %% 2 == 1){
    positive_frequencies <- runif((window_size + 1)/2, min = -pi, max = pi)
    phase <- c( positive_frequencies, -rev( positive_frequencies[-1] ) )
  }else{
    positive_frequencies <- runif((window_size/2) + 1, min = -pi, max = pi)
    phase <- c( positive_frequencies, -rev( positive_frequencies[ 2:(window_size/2) ] ) )
  }
  
  return(phase)
}

phase_from_partial <- function(partial_signal, partial_scaling, 
  window_function = hamming(length(partial_signal))){
  
  if(!is.vector(partial_signal)) stop('partial_signal must be a vector')
  if(!is.vector(partial_scaling)) stop('partial_scaling must be a vector')
  if(length(partial_signal) != length(partial_scaling)){
    stop('partial_signal and partial_scaling must be the same length')
  }
  
  scaled_signal <- partial_signal / partial_scaling
  scaled_signal[is.nan(scaled_signal)] <-  0
  
  partial_fft <- fft(window_function*scaled_signal)
  
  phase <- atan2(Im(partial_fft), Re(partial_fft))
}

rtisi_la <- function(mag_stft, step_size, window_function = hamming(nrow(mag_stft)), 
  iterations = 10, fft_normaliser = nrow(mag_stft)){
  if(!is.matrix(mag_stft)) stop('mag_stft must be a matrix')
  if(is.complex(mag_stft)) stop('mag_stft must be a real valued matrix')
  
  window_size <- nrow(mag_stft)
  total_frames <- ncol(mag_stft)

  committed_signal <- rep(0,  step_size*(total_frames - 1) + window_size)
  committed_weight <- rep(0,  step_size*(total_frames - 1) + window_size)
  
  for(i in seq.int(total_frames)){
    look_ahead <- min((window_size / step_size) - 1, total_frames - i)
    in_frame_signal <- 
      committed_signal[(i-1)*step_size + (1:(window_size + look_ahead*step_size))]
    in_frame_weight <- 
      committed_weight[(i-1)*step_size + (1:(window_size + look_ahead*step_size))]
    
    message('Initial Iteration')
    
    for(j in 0:look_ahead){
      if(j == 0 & i == 1){
        
        message(paste('Generate a random initial phase for frame', 
          i, 'of magnitude spectrum'))
        initial_phase <- random_phase(window_size)
        
      }else 
        if(j == 0 & i != 1){
        
        message(
          paste(
          'Committed signal used to generate inital phase from partial signal for frame', 
            i
            )
          )
        initial_phase <- phase_from_partial(
          partial_signal = committed_signal[j*step_size + 1:window_size], 
          partial_scaling = committed_weight[j*step_size + 1:window_size], 
          window_function = window_function
          )
        
      }else{

        initial_phase <- phase_from_partial(
          partial_signal = in_frame_signal[j*step_size + 1:window_size], 
          partial_scaling =in_frame_weight[j*step_size + 1:window_size], 
          window_function = window_function
          )
        
      }
      
      in_frame_signal[j*step_size + 1:window_size] <- 
        in_frame_signal[j*step_size + 1:window_size] + 
        (Re(fft(mag_stft[, i + j]*exp((0+1i)*initial_phase), inverse = T)) / 
            fft_normaliser)
    
      in_frame_weight[j*step_size + 1:window_size] <- 
        in_frame_weight[j*step_size + 1:window_size] + 
        window_function
    }
    
    frame_i_estimate <- (in_frame_signal / in_frame_weight)[1:window_size]
    
    temp_stft <- complex_stft(frame_i_estimate, window_size = window_size, 
      step_size = step_size)
    
    estimated_phase <- atan2(Im(temp_stft), Re(temp_stft))
    
    message(paste('Iteratively improving phase estimates for frame', i))
    for(k in seq.int(iterations)){
      in_frame_signal <- 
        committed_signal[(i-1)*step_size + (1:(window_size + look_ahead*step_size))]
      in_frame_weight <- 
        committed_weight[(i-1)*step_size + (1:(window_size + look_ahead*step_size))]
      
      for(j in 0:look_ahead){
        if(j == 0){
          
          initial_phase <- estimated_phase
          
        }else if(j > 0){
          
          initial_phase <- phase_from_partial(
            partial_signal = in_frame_signal[j*step_size + 1:window_size], 
            partial_scaling =in_frame_weight[j*step_size + 1:window_size], 
            window_function = window_function)
          
        }
        
        in_frame_signal[j*step_size + 1:window_size] <- 
          in_frame_signal[j*step_size + 1:window_size] + 
          (Re(fft(mag_stft[, i + j]*exp((0+1i)*initial_phase), inverse = T)) / 
              fft_normaliser)
        
        in_frame_weight[j*step_size + 1:window_size] <- 
          in_frame_weight[j*step_size + 1:window_size] + 
          window_function
      }
      
      frame_i_estimate <- (in_frame_signal / in_frame_weight)[1:window_size]
      
      temp_stft <- complex_stft(frame_i_estimate, window_size = window_size, 
        step_size = step_size)
      
      estimated_phase <- atan2(Im(temp_stft), Re(temp_stft))
    }
    
    message(paste('Committing frame', i))
    
    committed_signal[(i-1)*step_size + 1:window_size] <- 
      committed_signal[(i-1)*step_size + 1:window_size] + 
      (Re(fft(mag_stft[, i]*exp((0+1i)*estimated_phase), inverse = T)) / 
          fft_normaliser)*
      window_function
    
    committed_weight[(i-1)*step_size + 1:window_size] <- 
      committed_weight[(i-1)*step_size + 1:window_size] +
      window_function
  }
  
  reconstructed_signal <- committed_signal / committed_weight
  return(reconstructed_signal)
}
```

The RTISI-LA algorithm estimates the time domain signal that produced a given magnitude stft spectrum. Thus in order to invert a preprocessed spectrogram it must first be converted into the format of a magnitude STFT spectrum.

```{r smoothed surfaces to magnitude spectra, echo = T}
surface_to_mag_stft <- function(surface, surface_freq, pos_mag_freq){
  mag_surface <- sqrt(10^(surface / 10))
  
  mag_stft <- matrix(nrow = length(pos_mag_freq), ncol = ncol(surface))
  mag_stft[
    pos_mag_freq < (min(surface_freq) - 0.1) | 
      pos_mag_freq > (max(surface_freq) + 0.1)
    ] <- 0
  for(i in seq.int(ncol(surface))){
    
    mag_stft[is.na(mag_stft[,i]),i] <-
      sqrt(10^((
        spline(surface_freq, surface[,i], xout = pos_mag_freq[is.na(mag_stft[,i])])$y
        )/10))
    
  }
  
  mag_stft <- rbind(mag_stft, mag_stft[(length(pos_mag_freq) - 1):2,])
  
  return(mag_stft)
}
```

```{r data}
sp <- sdsBAT::phylogeny$tip.label

df <- readRDS('preprocessed_calls.RDS') %>% 
  select(bat, species, family, full) %>% 
  mutate(species = factor(species, levels = sp))

t <- seq(from = 0, to = 1, length.out = 100) %>% 
  extract(c(F, T))

f <- seq(from = 0, to = 250, length.out = 257)
restricted_f <- f %>% 
  extract(f > 9 & f < 212) %>% 
  extract(c(F, T))

remove(sp, f)
```

```{r mean spectrograms}
temp <- df %>% 
  group_by(bat, family) %>% 
  summarise(species = unique(species))

bat <- temp %>% use_series(bat)
species <- temp %>% use_series(species)
family <- temp %>% use_series(family)

by_bat <- df %>% 
  use_series(full) %>% 
  unlist %>% 
  array(dim = c(104*50, df %>% nrow)) %>%
  apply(1, function(x) tapply(x, df %>% use_series(bat), mean)) %>% 
  as_tibble() %>% 
  mutate(bat = bat, species = species, family = family) %>% 
  select(bat, species, family, starts_with('V')) %>% 
  group_by(bat, family, species) %>% 
  nest %>% 
  mutate(full = map(data, function(x) x %>% unlist %>% array(dim = c(104,50)))) %>% 
  select(-data) %>% 
  ungroup

temp <- by_bat %>% 
  group_by(species) %>% 
  summarise(family = unique(family))

species <- temp %>% use_series(species)
family <- temp %>% use_series(family)


by_species <- by_bat %>% 
  use_series(full) %>% 
  unlist %>% 
  array(dim = c(104*50, by_bat %>% nrow)) %>%
  apply(1, function(x) tapply(x, by_bat %>% use_series(species), mean)) %>% 
  as_tibble() %>% 
  mutate(species = species, family = family) %>% 
  select(species, family, starts_with('V')) %>% 
  group_by(family, species) %>% 
  nest %>% 
  mutate(full = map(data, function(x) x %>% unlist %>% array(dim = c(104,50)))) %>% 
  select(-data) %>% 
  ungroup

family <- by_species %>% use_series(family) %>% unique

by_family <- by_species %>% 
  use_series(full) %>% 
  unlist %>% 
  array(dim = c(104*50, by_species %>% nrow)) %>%
  apply(1, function(x) tapply(x, by_species %>% use_series(family), mean)) %>% 
  as_tibble() %>% 
  mutate(family = family) %>% 
  select(family, starts_with('V')) %>% 
  group_by(family) %>% 
  nest %>% 
  mutate(full = map(data, function(x) x %>% unlist %>% array(dim = c(104,50)))) %>% 
  select(-data) %>% 
  ungroup

rm(bat, family, species, temp)
```

```{r bootstrapped sample}
temp <- df %>% 
  group_by(bat) %>% 
  sample_n(1) %>% 
  ungroup %>% 
  group_by(species) %>% 
  sample_n(4) %>% 
  ungroup

boot_mean <- df %>% 
  group_by(species) %>% 
  summarise(family = unique(family)) %>% 
  ungroup %>% 
  cbind(
    check <- temp %>%
      use_series(full) %>% 
      unlist %>% 
      array(dim = c(104*50, temp %>% nrow)) %>% 
      apply(1, function(x) tapply(x, temp %>% use_series(species), mean))
  ) %>% 
  group_by(species, family) %>% 
  nest %>%
  mutate(full = map(data, function(x) x %>% unlist %>% array(dim = c(104,50)))) %>% 
  select(-data) %>% 
  ungroup

for(i in 1:24){
  temp <- df %>% 
    group_by(bat) %>% 
    sample_n(1) %>% 
    ungroup %>% 
    group_by(species) %>% 
    sample_n(4) %>% 
    ungroup
  
  temp_mean <- df %>% 
  group_by(species) %>% 
  summarise(family = unique(family)) %>% 
  ungroup %>% 
  cbind(
    check <- temp %>%
      use_series(full) %>% 
      unlist %>% 
      array(dim = c(104*50, temp %>% nrow)) %>% 
      apply(1, function(x) tapply(x, temp %>% use_series(species), mean))
  ) %>% 
  group_by(species, family) %>% 
  nest %>%
  mutate(full = map(data, function(x) x %>% unlist %>% array(dim = c(104,50)))) %>% 
  select(-data) %>% 
  ungroup
  
  boot_mean <- rbind(boot_mean, temp_mean)
  remove(temp, temp_mean)
}

boot_mean <- boot_mean %>% 
  arrange(species)
```

```{r bootstrapped sample tree}
n_samples <- 25
boot_tree <- phylogeny

for(i in phylogeny %>% use_series(tip.label)){
  individuals <- stree(n_samples)
  individuals$edge.length <- rep(0, n_samples)
  individuals$tip.label <- rep(i, n_samples)
  
  boot_tree <- bind.tree(boot_tree, individuals, which(boot_tree$tip.label == i))
}

rm(individuals, n_samples)
```

```{r pca}
pca <- boot_mean %>% 
  use_series(full) %>%
  unlist %>% 
  array(dim = c(50*104, df %>% nrow)) %>% 
  extract(, sample.int(df %>% nrow)) %>% 
  t %>% 
  prcomp
```

```{r centre spectrograms}
boot_mean <- boot_mean %>% 
  mutate(centred = map(full, function(x) x %>% subtract(pca %>% use_series(center))))
```

```{r pc scores}
n <- 15

pc_scores <- boot_mean %>% 
  select(species, family) %>% 
  cbind(pca %>% 
    use_series(rotation) %>% 
    extract(, 1:n) %>% 
    t %>% 
    multiply_by_matrix( boot_mean %>%
        select(centred) %>%
        unlist %>% 
        array(dim = c(104*50, boot_mean %>% nrow))
    ) %>% 
    apply(2, divide_by, pca %>% use_series(sdev) %>% extract(1:n)) %>% 
    t %>% 
    as_tibble
  )
```

```{r pc hyperparameters}
pc_mle <- pc_scores %>% 
  select(starts_with('PC')) %>%
  apply(2, sdsBAT:::pou_type2mle, boot_tree, logl_function = pou_logl_fast,
  optim_function = "optim", optim_method = "Nelder-Mead",
  lower_initialisation = c(0, 0, 0), upper_initialisation = c(1, 100, 1),
  n_restarts = 5)

rownames(pc_mle) <- c('s_p', 'l', 's_n', 'logl')

print('BOOTSTRAPPED PC MLE HYP')
pc_mle %>% round(2)
```

```{r project species level}
by_species <- by_species %>% 
  mutate(centred = map(full, function(x) x %>% subtract(pca %>% use_series(center))))

n <- 15

by_species_proj <- by_species %>% 
  select(species) %>% 
  cbind(
    pca %>%
      use_series(rotation) %>% 
      extract(, 1:n) %>% 
      t %>% 
      multiply_by_matrix( by_species %>%
          select(centred) %>%
          unlist %>% 
          array(dim = c(104*50, by_species %>% nrow))
      ) %>% 
      t %>% 
      multiply_by_matrix(
        pca %>% 
          use_series(rotation) %>% 
          extract(, 1:n) %>% 
          t
      ) %>% 
      as_tibble()
  ) %>% 
  group_by(species) %>% 
  nest %>% 
  mutate(projection = map(data, function(x){
    x %>% unlist %>% add(pca %>% use_series(center)) %>% array(dim = c(104, 50))
  })) %>% 
  select(-data)
```

When comparing the spectrograms there could be magnitude and scale differences between spectrograms, which do not affect our interpretation of the spectrogram but which do lead to reconstruction errors. With this in mind, spectrograms will be rescaled to lie on the interval $[0,1]$ before calculating error metrics.

```{r error metric}
mae <- function(x, y){
  x %>% subtract(y) %>% abs %>% mean
}

scaled_mae <- function(x, y){
  new_x <- x %>% subtract(x %>% min)
  new_x <- new_x %>% divide_by(new_x %>% max)
  
  new_y <- y %>% subtract(y %>% min)
  new_y <- new_y %>% divide_by(new_y %>% max)
  
  new_x %>% subtract(new_y) %>% abs %>% mean
}

by_species_proj <- by_species_proj %>% 
  mutate(scaled_mae = NA, mae = NA)

for(i in by_species_proj %>% nrow %>% seq.int){
  by_species_proj[i, 'scaled_mae'] <- scaled_mae(by_species[i, 'full'][[1]][[1]], by_species_proj[i, 'projection'][[1]][[1]])
  by_species_proj[i, 'mae'] <- mae(by_species[i, 'full'][[1]][[1]], by_species_proj[i, 'projection'][[1]][[1]])
}

#by_species_proj %>% use_series(mae)
#by_species_proj %>% use_series(scaled_mae)

```

```{r loso estimates}
ou_K <- function(distance_matrix, ln_hyperparameters){
  sp <- exp(2*ln_hyperparameters[1])
  l <- exp(ln_hyperparameters[2])
  sn <- exp(2*ln_hyperparameters[3])
  
  K <- distance_matrix %>% 
    abs %>%
    multiply_by(-1) %>% 
    divide_by(l) %>% 
    exp %>%
    multiply_by(sp) %>% 
    add(distance_matrix %>% nrow %>% diag %>% multiply_by(sn))
  
  return(K)
}

sp <- by_species %>% use_series(species)
pc <- colnames(pc_mle)

D <- cophenetic.phylo(boot_tree)

sp_map <- matrix(nrow = sp %>% length, ncol = pc %>% length) %>% 
  set_rownames(sp) %>% 
  set_colnames(pc)

sp_post_var <- matrix(nrow = sp %>% length, ncol = pc %>% length) %>% 
  set_rownames(sp) %>% 
  set_colnames(pc)

for(i in sp){
  for(j in pc){
    K <- ou_K(D, pc_mle %>% extract(1:3, j) %>%  log)
    inv_K_obs <- K[rownames(K) != i, colnames(K) != i] %>% solve
    
    sp_map[i, j] <- K[i %>% as.character, colnames(K) != i] %>% 
      multiply_by_matrix(inv_K_obs) %>% 
      multiply_by_matrix(pc_scores %>% filter(species != i) %>% extract2(j))
    
    sp_post_var[i, j] <- K[i %>% as.character, i %>% as.character] %>%
      subtract(
        K[i %>% as.character, colnames(K) != i] %>% 
          multiply_by_matrix(inv_K_obs) %>%
          multiply_by_matrix(K[colnames(K) != i, i %>% as.character])
        )
  }
}
```

```{r spectrogram reconstruction}
loso_recon <- sp_map %>% 
  multiply_by_matrix(
    pca %>% 
      use_series(rotation) %>% 
      extract(,1:15) %>% 
      apply(1, multiply_by, pca %>% use_series(sdev) %>% extract(1:15))
  ) %>% 
  as_tibble() %>% 
  cbind(by_species %>% select(species, family), .) %>% 
  group_by(species, family) %>% 
  nest %>%
  mutate(centred = map(data, function(x) x %>% unlist %>% array(dim = c(104,50)))) %>% 
  select(-data) %>% 
  ungroup %>% 
  mutate(full = map(centred, function(x) x %>% add(pca %>% use_series(center))))

loso_recon <- loso_recon %>% 
  mutate(mae = NA, scaled_mae = NA)

for(i in loso_recon %>% nrow %>% seq.int){
  loso_recon[i, 'mae'] <- mae(by_species[i, 'full'][[1]][[1]], loso_recon[i, 'full'][[1]][[1]])
  loso_recon[i, 'scaled_mae'] <- scaled_mae(by_species[i, 'full'][[1]][[1]], loso_recon[i, 'full'][[1]][[1]])
}
```


```{r}
order <- phylogeny %>% 
  fortify %>% 
  subset(isTip) %>% 
  with(label[order(y, decreasing=T)])

index <- sapply(loso_recon$species, function(x) which(order == x))

order <- factor(order, levels = rev(order))

plot_a <- ggtree(phylogeny) +
  geom_tiplab(hjust = 0.5, offset = 2) +
  theme(plot.margin = margin(1, 0, 0.75, 0, "cm")) +
  labs(
    title = 'Bat Phylogenetic Tree'
  )

plot_b <- loso_recon %>%
  mutate(refactored_species = order[index]) %>% 
  ggplot() +
  geom_point(aes(x = mae, y = refactored_species, color = family)) +
  theme(
    axis.text.y = element_blank(), 
    axis.ticks.y = element_blank(),
    legend.position = 'none') +
  labs(
    title = 'Mean Absolute Error\nof LOSO-CV\nReconstruction',
    x = 'Absolute Error',
    y = ''
  )

plot_c <- loso_recon %>%
  mutate(refactored_species = order[index]) %>% 
  ggplot() +
  geom_point(aes(x = scaled_mae, y = refactored_species, color = family)) +
  labs(
    title = 'Scaled Mean Absolute\nError of LOSO-CV\nReconstruction',
    x = 'Scaled Absolute Error',
    y = ''
  )

gridExtra::grid.arrange(plot_a, plot_c, ncol = 2)
gridExtra::grid.arrange(plot_a, plot_b, plot_c, ncol=3)
#cowplot::plot_grid(plot_a, plot_b, labels=c('A', 'B'))
remove(plot_a, plot_b, plot_c)

index %>% c

order[index] == by_species_proj$species
```

```{r ancestral reconstruction}
pc <- colnames(pc_mle)
D <- dist.nodes(boot_tree)

anc <- colnames(D) %>% extract(-550:-1)

ancestral_map <- matrix(nrow = anc %>% length, ncol = pc %>% length) %>% 
  set_rownames(anc) %>% 
  set_colnames(pc)

ancestral_post_var <- matrix(nrow = anc %>% length, ncol = pc %>% length) %>% 
  set_rownames(anc) %>% 
  set_colnames(pc)

for(i in anc){
  for(j in pc){
    K <- ou_K(D, pc_mle %>% extract(1:3, j) %>%  log)
    inv_K_obs <- K[!(rownames(K) %in% anc), !(colnames(K) %in% anc)] %>% solve
    
    ancestral_map[i, j] <- K[i %>% as.character, !(colnames(K) %in% anc)] %>% 
      multiply_by_matrix(inv_K_obs) %>% 
      multiply_by_matrix(pc_scores %>% extract2(j))
    
    ancestral_post_var[i, j] <- K[i %>% as.character, i %>% as.character] %>%
      subtract(
        K[i %>% as.character, !(colnames(K) %in% anc)] %>% 
          multiply_by_matrix(inv_K_obs) %>%
          multiply_by_matrix(K[!(colnames(K) %in% anc), i %>% as.character])
        )
  }
}

ancestor <-  c('23', '24','25', '26', '27', 'Arja', '28', 'Stli', 'Stlu', '29', '30', 'Dero', 'Maca', 'Leye', '31', 'Mome', '32', 'Ptpa', '33', 'Ptda', 'Ptpe', '34', '35', '36', '37', '38', 'Anpa', 'Pihe', 'Epfu', 'Idph', '39', '40', 'Laci', '41', 'Labl', 'Laxa', '42', 'Myvo', 'Myyu', '43', 'Tabr', 'Nyfe', 'Bapl' ) %>% factor

ancestral_recon <- ancestral_map %>% 
  multiply_by_matrix(
    pca %>% 
      use_series(rotation) %>% 
      extract(,1:15) %>% 
      apply(1, multiply_by, pca %>% use_series(sdev) %>% extract(1:15))
  ) %>% 
  as_tibble() %>% 
  cbind(ancestor, .) %>% 
  group_by(ancestor) %>% 
  nest %>%
  mutate(centred = map(data, function(x) x %>% unlist %>% array(dim = c(104,50)))) %>% 
  select(-data) %>% 
  ungroup %>% 
  mutate(full = map(centred, function(x) x %>% add(pca %>% use_series(center))))
```

```{r}
plot.phylo(phylogeny)
nodelabels()
```

```{r}
check = 'Stlu'
expand.grid(freq = restricted_f, time = t) %>% 
  mutate(spec = ancestral_recon %>% filter(ancestor == check) %>% use_series(full) %>% unlist) %>% 
  ggplot() +
  geom_raster(aes(x = time, y = freq, fill = spec), show.legend = F) +
  viridis::scale_fill_viridis(option = 'D') +
  theme_classic() +
  scale_x_continuous(breaks = c(0, 0.5, 1)) +
  labs(
    title = 'Ancestral Spectrogram' %>% paste(check, .),
    x = 'Time (Absolute)',
    y = 'Frequency (kHz)'
  )
```

```{r invert complex stft, echo = T}
inv_complex_stft <- function(stft, window_size = nrow(stft), 
  window_function = hamming(window_size), 
  step_size = ceiling(window_size/2), stft_scale = nrow(stft)){
  
  if (!is.complex(stft) & !is.matrix(stft)) stop("'stft' has to be a complex matrix.")
  
  if (length(window_size) > 1) stop("Cannot compute for multiple window sizes")
  
  window_offset <- seq(1, by = step_size, length.out = ncol(stft))
  
  inv_stft <- (Re(mvfft(stft, inverse = T))) / stft_scale
  
  signal <- matrix(0, nrow = max(window_offset) + window_size - 1, 
    ncol = length(window_offset))
  weight <-  matrix(0, nrow = max(window_offset) + window_size - 1, 
    ncol = length(window_offset))

  temp_w <- apply(array(1, dim = dim(inv_stft)),
    2, `*`, window_function)
  
  for(i in seq_along(window_offset)){
    signal[window_offset[i]:(window_offset[i] + window_size - 1), i] <- inv_stft[, i]
    weight[window_offset[i]:(window_offset[i] + window_size - 1), i] <- temp_w[, i]
  }
  
  weight <- rowSums(weight)
  signal <- rowSums(signal) / weight
  
  return(signal)
}
```



```{r}
library(audio)
mag_stft <- surface_to_mag_stft(ancestral_recon %>% filter(ancestor == check) %>% use_series(full) %>% extract2(1), restricted_f, seq(0, 250, length.out = 257))

signal <- rtisi_la(mag_stft, step_size = 2^5, window_function = hamming(nrow(mag_stft)), 
  iterations = 10, fft_normaliser = nrow(mag_stft))
  
plot(signal %>% divide_by(signal %>% abs %>% max), type = 'l')

sig <- audio::audioSample(signal %>% divide_by(signal %>% abs %>% max) %>% c, rate = 50000)
play(signal %>% divide_by(signal %>% abs %>% max) %>% c, rate = 5000)

```

