---
title: "Port Preprocessed Data to R"
author: "J.P. Meagher"
date: "20 October 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F)
```

As MATLAB offered a far superior toolkit for preprocessing the echolocation calls, this preprocessing was done in that environment. The preprocessed spectrograms produced must now be ported into R for further analysis.

# Packages

The key package here is the `R.matlab` package which allows the reading and writing of `mat` files.

The tidyverse and batwork packages are included as the spectrograms will be appended to the `mexican_bat_calls` dataset.

```{r packages}
library(tidyverse)
library(R.matlab)
library(batwork)
library(magrittr)
```

# Load Data

Convert the mat files to a list then extract the spectrograms into a list with a convenient format.

These columns will be concatenated with the `mexican_bat_calls` dataset and then saved for future reference.

```{r load r data}
df <- mexican_bat_calls
```

```{r add family variable}
fam <- c('Emb', 'Mol', 'Ves', 'Mor', 'Phy')

family_key <- df %>% 
  select(species) %>% 
  unique %>% 
  mutate(family = c(fam[3], fam[5], fam[1], fam[5], fam[3], fam[3], fam[3], fam[3], fam[3], fam[5],
  fam[5], fam[4], fam[3], fam[3], fam[2], fam[3], fam[4], fam[4], fam[4], fam[5], 
  fam[5], fam[2])) %>% 
  mutate(family = factor(family))

sp <- family_key %>% use_series(species)
fam <- family_key %>% use_series(family)

df <- df %>% 
  mutate(family = species %>% sapply(function(x) fam %>% extract(x %>% equals(sp) %>% which))) %>% 
  select(bat, species, family, sex, calls)

rm(fam, sp, family_key)
```

```{r load matlab data}
loc <- 'C:/Users/joe/Documents/MATLAB/ancestral-reconstruction'
full_fl <- 'preprocessed_full_spectrograms_reordered.mat'
thresh_fl <- 'preprocessed_thresholded_spectrograms_reordered.mat'

full <- readMat(paste(loc, full_fl, sep = '/')) %>%
  extract2(1) %>% 
  lapply(extract2, 1) %>% 
  lapply(extract, c(F, T), c(F, T))

thresholded <- readMat(paste(loc, thresh_fl, sep = '/')) %>%
  extract2(1) %>% 
  lapply(extract2, 1) %>% 
  lapply(extract, c(F, T), c(F, T))

df <- cbind(df, I(full), I(thresholded))

rm(full, full_fl, loc, thresh_fl, thresholded)
```

```{r save new df}
saveRDS(df, 'preprocessed_calls.RDS')
```



