---
title: "Reconstruction from Components"
author: "J.P. Meagher"
date: "31 October 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

It has been shown that the separability of spectrograms over time and frequency is not appropriate in this application and so I will proceed by investigating the accuracy of reconstructed species mean spectrograms from the components identified.

# Packages

```{r}
library(tidyverse)
library(magrittr)
library(sdsBAT)
```

# Data

```{r}
orig_df <- readRDS('preprocessed_calls.RDS')

df <- orig_df %>% 
  select(bat, species, full) %>% 
  mutate(flat = map(full, c))

by_species <- readRDS('by_species.RDS') %>% 
  select(species, full) %>% 
  mutate(flat = map(full, c)) 
  
```

# PCA

Run a Principal Components Analysis (PCA) on the sample of call spectrograms.

This requires pulling the spectrograms out of the dataframe in which they are stored, flattening them into vectors of length $104 \times 50$, concatenating these vectors into a single matrix such that the rows of this matrix represent different calls and then performing a standard PCA.

These Principal Components provide the basis from which Independent Components will be constructed. 

```{r}
pca <- df %>% 
  select(full) %>% 
  unnest %>% 
  use_series(full) %>% 
  array(dim = c(50*104, df %>% nrow)) %>% 
  t %>% 
  prcomp
```

# Linear Model of Independent Components

The first analysis performed will simply fit 7 independent components to the species mean echolocation calls and explore the accuracy of these reconstructions. 

## ICA

```{r}
n <- 7
ica <- cubica34(pca %>% use_series(rotation) %>% extract(, 1:n) %>% t)
component <- ica %>% use_series(y) %>% t

component <- component %>% divide_by(component %>% t %>% multiply_by_matrix(component) %>% diag %>% extract(1) %>% sqrt)
```

## Centre Mean Spectrograms

```{r}
by_species <- by_species %>% 
  mutate(centred = map(flat, subtract, pca %>% use_series(center)))
```

## Fit Linear Model

```{r}
ica_linear_model <- function(response){
  lm(response ~ component)
}

by_species <- by_species %>% 
  mutate(model = map(centred, ica_linear_model))

mod <- by_species %>% 
  select(species, model)

# struggles with 5, 7, 15, 19, 20, 21
n <- 22

mod %>% 
  filter(species == species %>% extract(n)) %>% 
  use_series(model) %>% 
  flatten %>% 
  extract2('fitted.values') %>% 
  matrix(nrow = 104, ncol = 50) %>% 
  image

by_species %>% 
  filter(species == species %>% extract(n)) %>% 
  use_series(centred) %>% 
  extract2(1) %>% 
  matrix(nrow = 104, ncol = 50) %>% 
  image
```

